Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-01-20 spie/_region_.tex
now zoomed images

\message{ !name(paper.tex) !offset(374) }
\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
parts of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, RANSAC
\cite{fischler1981random} is again used to estimate the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. We use a 10 pixel threshold for RANSAC, so that SIFT matches are
labeled as outliers if their distance is not within 10 pixels of the
sampled average distance.

We now use the $d^x_{i,j}$ and $d^y_{i,j}$ distances between each pair
of images to refine their positions using weighted linear least
squares. The variables we wish to solve for are the $x$ and $y$
positions of our images, while our equations are the SIFT-based
distances between pairs of images, images fixed to geometry with 0 or
1 degrees of freedom, and the original camera poses. Ignoring
weighting for now, an example setup for solving
$\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $
with 3 images is shown below.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  0 & -m & 0 & 0 & 1 & 0\\
\end{pmatrix}
\]

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & y_1, & y_2, & y_3
\end{pmatrix}
\]

\[
\vec{\gamma}^T =
\begin{pmatrix}
  dx_{1,2}, &dy_{1,2}, &dx_{2,3}, &dy_{2,3}, &t1_x, &t1_y, & -mx_2 +
  ty_2
\end{pmatrix}
\]

In this scenario, a SIFT-based distance of $dx_{1,2}$, $dy_{1,2}$ was
calculated between images 1 and 2. This corresponds to the first and
second of $A$, while the third and fourth row of $A$ represent the
same for images 2 and 3. Rows 5 and 6 of $A$ correspond to a fixed
location of $tx_1$, $ty_1$ for image 1, while Row 7 corresponds to a
constraint to a line of slope $m$, with current location $tx_2$,
$ty_2$, both results of the geometry alignment procedure in Section
\ref{sec:geometryAlignment}. If we do not have enough SIFT matches, or
lack images matched to geometry, our matrix becomes rank-deficient and
our problem cannot be solved. As a result, we add rows for each image
corresponding to the original noisy image locations, and downweight
them heavily, e.g. with a weight of 0.01 versus 1.

Because our problem is linear, it is quickly solved, and after
applying the resulting shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
repeating features are still visible.

\message{ !name(paper.tex) !offset(667) }

>>>>

\message{ !name(paper.tex) !offset(371) }
\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align images by searching for corresponding points
between all pairs of overlapping images. We use feature alignment
rather than pixel or intensity-based alignment due to the differences
in lighting as well as possible occlusion among our images, both of
which feature alignment is less sensitive to \cite{lowe1999object,
  mikolajczyk2005performance, szeliski2006image}.  We use SiftGPU
\cite{siftgpu} for its high performance on both feature detection as
well as pairwise matching. These matches determine $dx$ and $dy$
distances between each pair of features for two image projections,
though these distances may not always be the same for different
features. Since indoor environments often contain repetitive features
such as floor tiles or doors, we need to ensure that SIFT-based
distances are reliable. First, we only perform alignment on parts of
images that overlap given the original noisy poses. Second, we discard
feature matches that correspond to an image distance greater than 200
mm from what the noisy poses estimate. In order to utilize the
remaining feature matches robustly, RANSAC \cite{fischler1981random}
is again used to estimate the optimal $dx_{i,j}$ and $dy_{i,j}$
distances between two images $i$ and $j$. We use a 5 mm threshold for
RANSAC, so that SIFT matches are labeled as outliers if their distance
is not within 5 mm of the sampled average distance.

We now use the $dx_{i,j}$ and $dy_{i,j}$ distances between each pair
of images to refine their positions using weighted linear least
squares. The variables we wish to solve for are the $_ix$ and $y_i$
positions of our images, while our equations are the SIFT-based
distances between pairs of images, images fixed to geometry with 0 or
1 degrees of freedom, and the original camera poses. Ignoring
weighting and the original camera poses for now, an example setup for
solving $\textrm{min}_{\vec{\beta}} ||A \vec{\beta} -
\vec{\gamma}||_2^2 $ with 3 images is shown below.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  0 & -m & 0 & 0 & 1 & 0\\
\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  x_1, \\ x_2, \\ x_3, \\ y_1, \\ y_2, \\ y_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  dx_{1,2}, \\ dy_{1,2}, \\ dx_{2,3}, \\ dy_{2,3}, \\ t1_x, \\ t1_y, \\ -mx_2 +
  ty_2
\end{pmatrix}
\]

In this scenario, a SIFT-based distance of $dx_{1,2}$, $dy_{1,2}$ was
calculated between images 1 and 2. This corresponds to the first and
second row of $A$, while the third and fourth row of $A$ represent the
same for images 2 and 3. Rows 5 and 6 of $A$ correspond to a fixed
location of $tx_1$, $ty_1$ for image 1, while Row 7 corresponds to a
constraint to a line of slope $m$, with current location $tx_2$,
$ty_2$, both possible results of the geometry alignment procedure in
Section \ref{sec:geometryAlignment}. If we do not have enough SIFT
matches, or lack images matched to geometry, our matrix becomes
rank-deficient and our problem cannot be solved. As a result, we add
rows for each image corresponding to the original noisy image
locations, and downweight them heavily, e.g. with a weight of 0.01
versus 1.

Because our problem is linear, it is easily and quickly solved, and
after applying the resulting shifts, our images overlap and match each
other with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
parallax effects are still visible.

\message{ !name(paper.tex) !offset(668) }

### 2013-01-20 spie/paper.tex
now zoomed images

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geomAlign}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) In this example, we project two images at once for
    the sake of demonstration, where the red lines are geometry-based
    lines where surfaces intersect, and the green ines are lines
    detected in the image via Hough transform. Above are the original
    projections using input camera poses, and below are the
    projections after rotation and translation for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

@@

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  0 & -m & 0 & 0 & 1 & 0\\
\end{pmatrix}
\]

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & y_1, & y_2, & y_3
\end{pmatrix}
\]

\[
\vec{\gamma}^T =
\begin{pmatrix}
  dx_{1,2}, &dy_{1,2}, &dx_{2,3}, &dy_{2,3}, &t1_x, &t1_y, & -mx_2 +
  ty_2
\end{pmatrix}
\]

>>>>

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=3in]{detectedLines.png}}
  \hspace{0.4 cm}
  \subfloat[][]{\includegraphics[height=3in, width=4in]{bothunalignedaligned.png}}
  \caption{(a) Lines detected in image projections via Hough transform
    (b) In this example, we project two images at once for the sake of
    demonstration, where the red lines are geometry-based lines where
    surfaces intersect, and the green ines are lines detected in the
    image. Above are the original projections using input camera
    poses, and below are the projections after rotation and
    translation for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

@@

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  0 & -m & 0 & 0 & 1 & 0\\
\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  x_1, \\ x_2, \\ x_3, \\ y_1, \\ y_2, \\ y_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  dx_{1,2}, \\ dy_{1,2}, \\ dx_{2,3}, \\ dy_{2,3}, \\ t1_x, \\ t1_y, \\ -mx_2 +
  ty_2
\end{pmatrix}
\]
